---
title: "FSA to HCC code"
author: "PO Couture"
format: html
server: shiny
---

## Shiny Documents

This Quarto document is made interactive using Shiny. Interactive documents allow readers to modify parameters and see the results immediately. Learn more about Shiny interactive documents at <https://quarto.org/docs/interactive/shiny/>.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change. This demonstrates how a standard R plot can be made interactive:

```{r}
library(sf)
library(dplyr)
library(ggplot2)

# ---- 1. HCC: NBHC 33 community boundaries (NB Open Data, GeoJSON) ----
hcc_url <- "https://gnb.socrata.com/api/geospatial/sdyn-7de6?method=export&format=GeoJSON"
hcc <- st_read(hcc_url)

# ---- 2. FSA: New Brunswick forward sortation areas (StatsCan ArcGIS REST) ----
#    PRUID 13 = New Brunswick, so this returns only NB FSAs, no national download.
fsa_url <- paste0(
  "https://geo.statcan.gc.ca/geo_wa/rest/services/2021/",
  "Cartographic_boundary_files/MapServer/14/query",
  "?where=", URLencode("PRUID='13'", reserved = TRUE),
  "&outFields=CFSAUID,PRUID&returnGeometry=true&outSR=3347&f=geojson"
)
fsa <- st_read(fsa_url)

# ---- 3. Common projection (StatsCan Lambert) before any overlay or area maths ----
hcc <- st_transform(hcc, 3347)
fsa <- st_transform(fsa, 3347)

# ---- 4. Overlay map ----
ggplot() +
  geom_sf(data = fsa, fill = NA, colour = "grey40", linewidth = 0.3) +
  geom_sf(data = hcc, fill = NA, colour = "firebrick", linewidth = 0.6) +
  theme_void()

# ---- 5. Quantified overlap: how each FSA splits across HCCs by area ----
#    Run names(hcc) first, then put the real community column where HCC_ID is.
overlap <- st_intersection(fsa, hcc) |>
  mutate(piece_area = as.numeric(st_area(geometry))) |>
  st_drop_geometry() |>
  group_by(CFSAUID) |>
  mutate(pct_of_fsa = 100 * piece_area / sum(piece_area)) |>
  ungroup() |>
  arrange(CFSAUID, desc(pct_of_fsa))
```


```{r}
library(sf); library(dplyr); library(leaflet); library(htmlwidgets)

# ---- cross-reference lists from overlap ----
fsa_to_hcc <- overlap |>
  group_by(CFSAUID) |>
  arrange(desc(pct_of_fsa), .by_group = TRUE) |>
  summarise(hcc_list = paste0(en_nom_com, " (", round(pct_of_fsa), "%)",
                              collapse = "<br>"), .groups = "drop")

hcc_to_fsa <- overlap |>
  group_by(com) |>
  arrange(desc(pct_of_fsa), .by_group = TRUE) |>
  summarise(fsa_list = paste0(CFSAUID, " (", round(pct_of_fsa), "%)",
                              collapse = "<br>"), .groups = "drop")

# ---- attach to geometries, project for leaflet ----
fsa_pop <- fsa |> left_join(fsa_to_hcc, by = "CFSAUID") |> st_transform(4326)
hcc_pop <- hcc |> left_join(hcc_to_fsa, by = "com")      |> st_transform(4326)
# sanity check: nrow(fsa_pop) should be > 0 and match your FSA count

pal <- colorFactor("Set2", domain = hcc_pop$nbhc_zone)
bb  <- as.numeric(st_bbox(fsa_pop))

# ---- map ----
m <- leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  fitBounds(bb[1], bb[2], bb[3], bb[4]) |>
  addPolygons(data = hcc_pop, group = "HCC",
              fillColor = ~pal(nbhc_zone), fillOpacity = 0.30,
              color = "#444444", weight = 2, opacity = 0.9,
              label = ~en_nom_com,
              popup = ~paste0("<b>HCC ", com, ": ", en_nom_com, "</b>",
                              "<br><b>Zone:</b> ", nbhc_zone,
                              "<br><b>FSAs inside:</b><br>", fsa_list),
              highlightOptions = highlightOptions(weight = 4, color = "#000000")) |>
  addPolygons(data = fsa_pop, group = "FSA",
              fill = TRUE, fillOpacity = 0,          # invisible fill = whole area clickable
              color = "#1A1A1A", weight = 2.2, opacity = 0.9, dashArray = "4",
              label = ~CFSAUID,
              popup = ~paste0("<b>FSA ", CFSAUID, "</b>",
                              "<br><b>Connected HCC(s):</b><br>", hcc_list),
              highlightOptions = highlightOptions(weight = 4, color = "#E0591A",
                                                  bringToFront = TRUE)) |>
  addLayersControl(overlayGroups = c("FSA", "HCC"),
                   options = layersControlOptions(collapsed = FALSE))

m
saveWidget(m, "fsa_hcc_map_1.html", selfcontained = TRUE)
```



